

##Columns Explaination
1) ID = internal kickstarter id
2) name = name of project - A project is a finite work with a clear goal that you’d like to bring to life. Think albums, books, or films.
3) category = category
4) main_category = category of campaign
5) currency = currency used to support
6) deadline = deadline for crowdfunding
7) goal = fundraising goal - The funding goal is the amount of money that a creator needs to complete their project.
8) launched = launched date launched
9) pledged = amount pledged by "crowd"
10) state = Current condition the project is in
11) backers = number of backers
12) country = country pledged from
13) usd pledged = amount of money pledged

```{r}
#reads in data into KickStar
KickStar <- read.csv("/Users/karlhickel/Desktop/ks-projects-201612.csv")
#removing a few columns with null values
KickStar <- subset(KickStar, select = -c(14,15,16,17))
colSums(is.na(KickStar))
KickStar<- KickStar[!is.na(KickStar$name),]
colSums(is.na(KickStar))
#cleaning up category column
KickStar$category <- factor(KickStar$category, levels = c("Zines","Young Adult", "World Music", "Workshops" ,"Woodworking", "Webseries", "Webcomics", "Web","Weaving", "Wearables", "Video Games", "Video Art", "Video", "Vegan", "Typography", "Translations", "Thrillers", "Theater", "Textiles", "Television", "Technology", "Taxidermy", "Tabletop Games", "Stationery", "Spaces", "Space Exploration", "Sound", "Software", "Small Batch", "Shorts", "Sculpture", "Science Fiction", "Romance", "Rock", "Robots", "Restaurants", "Residencies", "Ready-to-wear", "Radio & Podcasts", "R&B", "Quilts", "Puzzles", "Punk", "Publishing", "Public Art", "Product Design", "Printing", "Print", "Pottery", "Pop", "Poetry", "Plays", "Playing Cards", "Places", "Photography", "Photobooks", "Photo", "Pet Fashion", "Periodicals", "Performances", "Performance Art", "People", "Painting", "Nonfiction", "Nature", "Narrative Film", "Musical", "Music Videos", "Music", "Movie Theaters", "Mobile Games", "Mixed Media", "Metal", "Makerspaces", "Live Games", "Literary Journals", "Letterpress", "Latin", "Knitting", "Kids", "Journalism", "Jewelry", "Jazz", "Interactive Design", "Installations", "Indie Rock", "Immersive", "Illustration", "Horror", "Hip-Hop", "Hardware", "Graphic Novels", "Graphic Design", "Glass", "Gaming Hardware", "Games", "Gadgets", "Footwear", "Food Trucks", "Food", "Flight", "Fine Art", "Film & Video", "Fiction", "Festivals", "Fashion", "Farms", "Farmer's Markets", "Fantasy", "Family", "Faith", "Fabrication Tools", "Experimental", "Events", "Embroidery", "Electronic Music", "Drinks", "Drama", "Documentary", "DIY Electronics", "DIY", "Digital Art", "Design", "Dance", "Crochet", "Crafts", "Couture", "Country & Folk", "Cookbooks", "Conceptual Art", "Community Gardens", "Comics", "Comic Books", "Comedy", "Classical Music", "Civic Design", "Chiptune", "Childrenswear", "Children's Books", "Ceramics", "Candles", "Camera Equipment", "Calendars", "Blues", "Bacon", "Audio", "Art Books", "Art", "Architecture", "Apps", "Apparel", "Anthologies", "Animation", "Animals", "Action", "Accessories", "Academic", "3D Printing"))
colSums(is.na(KickStar))
KickStar<- KickStar[!is.na(KickStar$category),]
colSums(is.na(KickStar))
#cleaning main_category column
table(KickStar$main_category)
KickStar$main_category <- factor(KickStar$main_category, levels = c("Theater", "Technology", "Publishing", "Photography", "Music", "Journalism", "Games", "Food", "Film & Video", "Fashion", "Design", "Dance", "Crafts", "Comics", "Art"))
colSums(is.na(KickStar))
#clearning currency column
table(KickStar$currency)
KickStar$currency <- factor(KickStar$currency, levels = c("USD", "SGD", "SEK", "NZD", "NOK", "MXN", "HKD", "GBP", "EUR", "DKK", "CHF", "CAD", "AUD"))
colSums(is.na(KickStar))
#cleaning goal column
KickStar$goal <- as.numeric(KickStar$goal)
colSums(is.na(KickStar))
#cleaning pledge column
KickStar$pledged <- as.numeric(KickStar$pledged)
#cleaning state column
table(KickStar$state)
KickStar$state <- factor(KickStar$state, levels = c("failed", "successful", "canceled", "live", "undefined", "suspended"))
colSums(is.na(KickStar))
#cleaning backers column
KickStar$backers <- as.numeric(KickStar$backers)
colSums(is.na(KickStar))
#cleaning country column
table(KickStar$country)
KickStar$country <- factor(KickStar$country, levels = c("US", "SG", "SE", "NZ", "NO", "NL", "MX", "LU", "IT", "IE", "HK", "GB", "FR", "ES", "DK", "DE", "CH", "CA", "BE", "AU", "AT"))
colSums(is.na(KickStar))
KickStar<- KickStar[!is.na(KickStar$country),]
colSums(is.na(KickStar))
#cleaning usd.pledged column
KickStar$usd.pledged <- as.numeric(KickStar$usd.pledged)
colSums(is.na(KickStar))
#gets summary of data
summary(KickStar)

```



```{r}
#install.packages("ggplot2")
library(ggplot2)
#ggplot of Goal Amount and Pledge Amount by state of company
ggplot(data = KickStar, aes(goal, pledged)) + geom_smooth() + facet_wrap(~state) + labs(title = "Smooth Plot of Goal vs Pledge Amount by State of Company") + labs(x = "Goal Amount", y = "Pledge Amount")
#ggplot of Goal Amount and Pledge Aamount by country
ggplot(data = KickStar, aes(goal, pledged)) + geom_smooth() + facet_wrap(~country) + labs(title = "Smooth Plot of Goal vs Pledge Amount by Country") + labs(x = "Goal Amount", y = "Pledge Amount")
#ggplot of bar graph with type of currency and the amount pledged 
ggplot(data = KickStar, aes(x = currency, y = usd.pledged)) + geom_bar(stat = "identity") + labs(x = "Type of Currency", y = "Amount of $ Pledged") + labs(title = "Amount Pledged by Currency")
##ggplot of bar graph of count of each categories
ggplot(data = KickStar, aes(category)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title = "Count of Categories", x = "Categories", y = "Count")
#ggplot of bar graph of count by state 
ggplot(data = KickStar, aes(state)) + geom_bar() + labs(title = "Counts per State")
```

```{r}
#beginning of Lasso, Ridge, ElasticNet
#install.packages("useful")
library(useful)
myFormula <- as.formula(binomState ~ category + main_category + currency + goal + pledged + backers + country + usd.pledged)
library(glmnet)
Xvar <- build.x(myFormula, KickTrain) 
Yvar <- build.y(myFormula, KickTrain)
LassoMod <- cv.glmnet(x = Xvar, y = Yvar, alpha = 1, nfolds = 10)
plot(LassoMod)


#Ridge Mod.
RidgeMod <- cv.glmnet(x = Xvar, y = Yvar, alpha = 0, nfolds = 10)
plot(RidgeMod)

#Elastic Net
ElasticNet1 <- cv.glmnet(x = Xvar, y = Yvar, alpha = .25, nfolds = 10)
ElasticNet2 <- cv.glmnet(x = Xvar, y = Yvar, alpha = .50, nfolds = 10)
ElasticNet3 <- cv.glmnet(x = Xvar, y = Yvar, alpha = .75, nfolds = 10)
summary(ElasticNet1)

#MSE
View(KickStar)
predictedTrain <- predict(portGradeTrain, data = portData2Train)
predictedTest <- predict(portGradeTest, data = portData2Test)
mseTrain <- mean((portData2Train$G3 - predictedTrain)^2)
mseTest <- mean((portData2Test$G3 - predictedTest)^2)
mseTrain
mseTest




```


```{r}
#Logistic Regression
KickStar$binomState <- ifelse(KickStar$state == "failed",0,1)
set.seed(1861)
trainSize <- 0.75
trainInd <- sample(1:nrow(KickStar), size = floor(nrow(KickStar) * trainSize))
KickTrain <- KickStar[trainInd, ]
KickTest <- KickStar[-trainInd, ]
logitFit <- glm(binomState ~ goal + pledged + backers + usd.pledged + main_category + currency + country, data = KickTrain, family = binomial)
summary(logitFit)
exp(logitFit$coefficients)
plot(logitFit)
```







